---
title: "CRiSp Use Case: [Title of use case]"
format: html
---

```{r}
#| label: setup

# remotes::install_github("CityRiverSpaces/CRiSp")
library(CRiSp)
```

## Introduction

Choose one use case from Day 1 and describe it in one paragraph, including the following information:

- the discipline or domain of application
- the research question, objective or assignment
- the choice for single- or cross-case analysis
- conceptual, data-related or analytical challenges

Choose one (or more) method(s) of analysis from Day 2 and describe why you chose it/them to answer your question or respond to the assignment.

### Location

Set location

```{r}
city_name <- "Bucharest"
river_name <- "Dâmbovița"
```

Get OSM data for visualisation

```{r}
bb <- get_osm_bb(city_name)
crs <- get_utm_zone(bb)
streets <- get_osm_streets(bb, crs)
railways <- get_osm_railways(bb, crs)
```

## Delineation

```{r}
cr <- delineate(city_name, river_name,
                segments = TRUE,
                riverspace = TRUE)
```

```{r}
plot(cr$corridor, lwd = 3, border = "orange")
plot(cr$riverspace, col = "green", border = NA,  add = TRUE)
plot(streets$geometry, lwd = 0.5, add = TRUE)
plot(railways$geometry, lwd = 0.5, add = TRUE)
plot(cr$corridor, add = TRUE, lwd = 3, border = "orange")
plot(cr$segments, add = TRUE, border = "orange")
```

## Analysis

- Load a layer or set of layers of interest
- Carry out some spatial analysis
- Discuss the implications of using the areas delineated by CRiSp for the analysis
